Online Multiclass Linear Classification with Bandit Feedback

### 2018-12-18 arxiv-version/main.tex
Reduce to rational approximation

\begin{theorem}[Polynomial approximation of intersection of halfspaces I]
\label{theorem:polynomial-approximation-1}
Let $v_1, v_2, \dots, v_m \in \R^d$ be vectors such that $\norm{v_1},
\norm{v_2}, \dots, \norm{v_m} \le 1$. Let $\gamma > 0$. There exists a
multivariate polynomial $p:\R^d \to \R$ such that
\begin{enumerate}
\item $p(x) \ge 1$ for all $\displaystyle x \in \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \ge \gamma \right\}$
\item $p(x) \le -1$ for all $\displaystyle x \in \bigcup_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\}$
\item $\deg(p) \le m (2 \left\lceil \frac{1}{2} \log(4m + 1) \right\rceil + 1) (2 \left \lceil \log(1/\gamma) \right \rceil + 1)$
\item $\norm{p} \le ???$
\end{enumerate}
\end{theorem}

\begin{theorem}[Polynomial approximation of intersection of halfspaces II]
\label{theorem:polynomial-approximation-2}
Let $v_1, v_2, \dots, v_m \in \R^d$ be vectors such that $\norm{v_1},
\norm{v_2}, \dots, \norm{v_m} \le 1$. Let $\gamma > 0$. There exists a
multivariate polynomial $p:\R^d \to \R$ such that
\begin{enumerate}
\item $p(x) \ge 1/2$ for all $\displaystyle x \in \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \ge \gamma \right\}$
\item $p(x) \le -1/2$ for all $\displaystyle x \in \bigcup_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\}$
\item $\displaystyle \deg(p) = \left\lceil \log_2(2m) \right\rceil \cdot \left\lceil \sqrt{\frac{1}{\gamma}} \right\rceil$
\item $\displaystyle \norm{p} \le \left( 188 \left\lceil \log_2(2m) \right\rceil \cdot \left\lceil \sqrt{\frac{1}{\gamma}} \right\rceil \right)^{\frac{1}{2} \left\lceil \log_2(2m) \right\rceil \cdot \left\lceil \sqrt{\frac{1}{\gamma}} \right\rceil}$
\end{enumerate}
\end{theorem}

@@

We define several univariate polynomials
\begin{align*}
P_n(z) & = (z - 1) \prod_{i=1}^n (z - 2^i)^2 && \text{for $n \ge 0$,} \\
A_{n,k}(z) & = (P_n(z))^k - (P_n(-z))^k && \text{for $n,k \ge 0$,} \\
B_{n,k}(z) & = - (P_n(z))^k - (P_n(-z))^k && \text{for $n,k \ge 0$.}
\end{align*}
Let $r = 2 \left\lceil \frac{1}{2} \log(4m + 1) \right\rceil + 1$ and $s = \left \lceil \log(1/\gamma) \right \rceil$.
We define the polynomial $p:\R^d \to \R$ as
$$
p(x) = \left[ \sum_{i=1}^m A_{s,r}\left( \frac{\ip{v_i}{x}}{\gamma} \right) \prod_{\substack{j ~:~ 1 \le j \le m \\ j \neq i}} B_{s,r} \left( \frac{\ip{v_j}{x}}{\gamma} \right) \right]
- \left(m - \frac{1}{2} \right) \prod_{j=1}^m B_{s,r} \left( \frac{\ip{v_j}{x}}{\gamma} \right) \; .
$$
Clearly, $\deg(P_n) = 2n+1$. Thus, $\deg(A_{n,k})$ and $\deg(B_{n,k})$ are at
most  $(2n+1)k$, and therefore $\deg(p) \le m (2s+1) r$.

@@

\begin{proposition}[Properties of $P_n$]
\label{proposition:properties-of-p-n}
\hspace{1cm} % Dummy space
\begin{enumerate}
\item If $z \in [0,1]$ then $P_n(-z) \le P_n(z) \le 0$.
\item If $z \in [1,2^n]$ then $0 \le 4P_n(z) \le -P_n(-z)$.
\item If $z \ge 0$ then $-P_n(-z) \ge 2^{n(n-1)}$.
\end{enumerate}
\end{proposition}

@@

The third part follows from
$$
-P_n(-z) = (1+z) \prod_{i=1}^n (z+2^i)^2 \ge \prod_{i=1}^n 2^{2i} = 2^{n(n-1)} \; .
$$
\end{proof}

\begin{proposition}[Properties of $S_{n,r}$ and $B_{n,r}$]
For any integer $n \ge 0$ and $r = 2 \left\lceil \frac{1}{2} \log(4m + 1) \right\rceil + 1$,
\begin{enumerate}
\item If $z \in [1,2^n]$ then $S_{n,r}(z) \in [1,1+\frac{1}{2m}]$.
\item If $z \in [-2^n, -1]$ then $S_{n,r} \in [-1-\frac{1}{2m}, -1]$.
\item If $z \in [-1,1]$ then $|S_{n,r}(z)| \le 1$.
\item If $z \in [-2^n,2^n]$ then $B_{n,r}(z) \ge \left(1 - \frac{1}{4m+1} \right) 2^{n(n-1)r}$.
\end{enumerate}
\end{proposition}

@@

\begin{enumerate}
\item Observe that $S_{n,r}(z)$ can be written as
$$
S_{n,r}(z) = \frac{\displaystyle 1 + \left( - \frac{P_n(z)}{P_n(-z)}\right)^r}{\displaystyle 1 - \left( - \frac{P_n(z)}{P_n(-z)}\right)^r} = \frac{1 + c}{1 - c}
$$
where $c = \left( - \frac{P_n(z)}{P_n(-z)}\right)^r$. Since $z \in [1,2^n]$, by
part 2 of~\autoref{proposition:properties-of-p-n}, $c \in [0,\frac{1}{4^r}]$. By
the definition of $r$ that means that $c \in [0,\frac{1}{4m+1}]$. Thus,
$S_{n,r}(z) = \frac{1+c}{1-c} \in [1,1 + \frac{1}{2m}]$.

@@

Case $z \in [0,1]$. Since $r$ is odd and $P_n(z)$ is non-positive,
$$
B_{n,r}(z) = - (P_n(z))^r + \left(- P_{n}(-z)\right)^r \ge \left(- P_{n}(-z)\right)^r \ge 2^{n(n-1)r} \ge 2^{n(n-1)r} \left( 1 - \frac{1}{4m+1} \right) \; .
$$
where the second last inequality follows from part 3 of \autoref{proposition:properties-of-p-n}.

Case $z \in [1,2^n]$. Since $r$ is odd,
$$
B_{n,r}(z) = \left(- P_{n}(-z)\right)^r \left(1 - \left( - \frac{P_n(z)}{P_n(-z)}\right)^r \right) = \left(- P_{n}(-z)\right)^r (1 - c)
$$
where $c = \left( - \frac{P_n(z)}{P_n(-z)}\right)^r$. Since $z \in [1,2^n]$, by
part 2 of~\autoref{proposition:properties-of-p-n}, $c \in [0,\frac{1}{4^r}]$. By
the definition of $r$ that means that $c \in [0,\frac{1}{4m+1}]$. Thus,
$$
B_{n,r}(z) \ge \left(- P_{n}(-z)\right)^r \left( 1 - \frac{1}{4m+1} \right) \ge 2^{n(n-1)r} \left( 1 - \frac{1}{4m+1} \right) \; .
$$
where the last inequality follows from part 3 of \autoref{proposition:properties-of-p-n}.
\end{enumerate}
\end{proof}

>>>>

\begin{theorem}[Polynomial approximation of intersection of halfspaces I]
\label{theorem:polynomial-approximation-1}
Let $v_1, v_2, \dots, v_m \in \R^d$ be vectors such that $\norm{v_1},
\norm{v_2}, \dots, \norm{v_m} \le 1$. Let $\gamma \in (0,1)$. There exists a
multivariate polynomial $p:\R^d \to \R$ such that
\begin{enumerate}
\item $\displaystyle p(x) \ge m^{\frac{1}{2} m \log^2(1/\gamma)}$
for all $\displaystyle x \in \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \ge \gamma \right\}$

\item $\displaystyle p(x) \le - m^{\frac{1}{2} m \log^2(1/\gamma)}$
for all $\displaystyle x \in \bigcup_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\}$

\item $\deg(p) \le m (2 \left\lceil \frac{1}{4} \log_2(4m + 1) \right\rceil + 1) (2 \left \lceil \log(1/\gamma) \right \rceil + 1)$
\item $\norm{p} \le ???$
\end{enumerate}
\end{theorem}

\begin{theorem}[Polynomial approximation of intersection of halfspaces II]
\label{theorem:polynomial-approximation-2}
Let $v_1, v_2, \dots, v_m \in \R^d$ be vectors such that $\norm{v_1},
\norm{v_2}, \dots, \norm{v_m} \le 1$. Let $\gamma \in (0,1)$. There exists a
multivariate polynomial $p:\R^d \to \R$ such that
\begin{enumerate}
\item $p(x) \ge 1/2$ for all $\displaystyle x \in \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \ge \gamma \right\}$
\item $p(x) \le -1/2$ for all $\displaystyle x \in \bigcup_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\}$
\item $\displaystyle \deg(p) = \left\lceil \log_2(2m) \right\rceil \cdot \left\lceil \sqrt{\frac{1}{\gamma}} \right\rceil$
\item $\displaystyle \norm{p} \le \left( 188 \left\lceil \log_2(2m) \right\rceil \cdot \left\lceil \sqrt{\frac{1}{\gamma}} \right\rceil \right)^{\frac{1}{2} \left\lceil \log_2(2m) \right\rceil \cdot \left\lceil \sqrt{\frac{1}{\gamma}} \right\rceil}$
\end{enumerate}
\end{theorem}

@@

We define several univariate polynomials
\begin{align*}
P_n(z) & = (z - 1) \prod_{i=1}^n (z - 2^i)^2 && \text{for $n \ge 0$,} \\
A_{n,k}(z) & = (P_n(z))^k - (P_n(-z))^k && \text{for $n,k \ge 0$,} \\
B_{n,k}(z) & = - (P_n(z))^k - (P_n(-z))^k && \text{for $n,k \ge 0$.}
\end{align*}
Let $r = 2 \left\lceil \frac{1}{4} \log_2(4m + 1) \right\rceil + 1$ and $s = \left \lceil \log(1/\gamma) \right \rceil$.
We define the polynomial $p:\R^d \to \R$ as
$$
p(x) = \left[ \sum_{i=1}^m A_{s,r}\left( \frac{\ip{v_i}{x}}{\gamma} \right) \prod_{\substack{j ~:~ 1 \le j \le m \\ j \neq i}} B_{s,r} \left( \frac{\ip{v_j}{x}}{\gamma} \right) \right]
- \left(m - \frac{1}{2} \right) \prod_{j=1}^m B_{s,r} \left( \frac{\ip{v_j}{x}}{\gamma} \right) \; .
$$
Clearly, $\deg(P_n) = 2n+1$. Thus, $\deg(A_{n,k})$ and $\deg(B_{n,k})$ are at
most  $(2n+1)k$, and therefore $\deg(p) \le m (2s+1) r$.

@@

\begin{proposition}[Properties of $P_n$]
\label{proposition:properties-of-p-n}
\hspace{1cm} % Dummy space
\begin{enumerate}
\item If $z \in [0,1]$ then $P_n(-z) \le P_n(z) \le 0$.
\item If $z \in [1,2^n]$ then $0 \le 4P_n(z) \le -P_n(-z)$.
\item If $z \ge 0$ then $-P_n(-z) \ge 2^{n(n+1)}$.
\end{enumerate}
\end{proposition}

@@

The third part follows from
$$
-P_n(-z) = (1+z) \prod_{i=1}^n (z+2^i)^2 \ge \prod_{i=1}^n 2^{2i} = 2^{n(n+1)} \; .
$$
\end{proof}

\begin{proposition}[Properties of $S_{n,r}$ and $B_{n,r}$]
\label{proposition:properties-of-s-n-r}
For any integer $n \ge 0$ and $r = 2 \left\lceil \frac{1}{4} \log_2(4m + 1) \right\rceil + 1$,
\begin{enumerate}
\item If $z \in [1,2^n]$ then $S_{n,r}(z) \in [1,1+\frac{1}{2m}]$.
\item If $z \in [-2^n, -1]$ then $S_{n,r} \in [-1-\frac{1}{2m}, -1]$.
\item If $z \in [-1,1]$ then $|S_{n,r}(z)| \le 1$.
\item If $z \in [-2^n,2^n]$ then $B_{n,r}(z) \ge \left(1 - \frac{1}{4m+1} \right) 2^{n(n+1)r}$.
\end{enumerate}
\end{proposition}

@@

\begin{enumerate}
\item Observe that $S_{n,r}(z)$ can be written as
$$
S_{n,r}(z) = \frac{\displaystyle 1 + \left( - \frac{P_n(z)}{P_n(-z)}\right)^r}{\displaystyle 1 - \left( - \frac{P_n(z)}{P_n(-z)}\right)^r} = \frac{1 + c}{1 - c}
$$
where $c = \left( - \frac{P_n(z)}{P_n(-z)}\right)^r$. Since $z \in [1,2^n]$, by
part 2 of~\autoref{proposition:properties-of-p-n}, $c \in [0,\frac{1}{4^r}]$. Since
$r \ge \frac{1}{2} \log_2(4m+1)$ that means that $c \in [0,\frac{1}{4m+1}]$. Thus,
$S_{n,r}(z) = \frac{1+c}{1-c} \in [1,1 + \frac{1}{2m}]$.

@@

Case $z \in [0,1]$. Since $r$ is odd and $P_n(z)$ is non-positive,
$$
B_{n,r}(z) = - (P_n(z))^r + \left(- P_{n}(-z)\right)^r \ge \left(- P_{n}(-z)\right)^r \ge 2^{n(n+1)r} \ge 2^{n(n+1)r} \left( 1 - \frac{1}{4m+1} \right) \; .
$$
where the second last inequality follows from part 3 of \autoref{proposition:properties-of-p-n}.

Case $z \in [1,2^n]$. Since $r$ is odd,
$$
B_{n,r}(z) = \left(- P_{n}(-z)\right)^r \left(1 - \left( - \frac{P_n(z)}{P_n(-z)}\right)^r \right) = \left(- P_{n}(-z)\right)^r (1 - c)
$$
where $c = \left( - \frac{P_n(z)}{P_n(-z)}\right)^r$. Since $z \in [1,2^n]$, by
part 2 of~\autoref{proposition:properties-of-p-n}, $c \in [0,\frac{1}{4^r}]$. By
the definition of $r$ that means that $c \in [0,\frac{1}{4m+1}]$. Thus,
$$
B_{n,r}(z) \ge \left(- P_{n}(-z)\right)^r \left( 1 - \frac{1}{4m+1} \right) \ge 2^{n(n+1)r} \left( 1 - \frac{1}{4m+1} \right) \; .
$$
where the last inequality follows from part 3 of \autoref{proposition:properties-of-p-n}.
\end{enumerate}
\end{proof}

To prove parts 1 and 2 of \autoref{theorem:polynomial-approximation-1}
note that \autoref{proposition:properties-of-s-n-r} implies that
\begin{align*}
|p(x)|
& = |Q(x)| \prod_{j=1}^m B_{s,r} \left( \frac{\ip{v_j}{x}}{\gamma} \right) \\
& \ge |Q(x)| \left( 2^{s(s+1)r} \left( 1 - \frac{1}{4m+1} \right) \right)^m \\
& \ge |Q(x)| \cdot 2^{s(s+1)rm} e^{-4} && \text{(since $\left(1-\frac{1}{4m+1}\right)^m \ge e^{-4}$)} \\
& \ge 4 |Q(x)| \cdot 2^{s^2rm} && \text{(since $e^{-4} \ge 1/2$ and $2^{rms} \ge 8$)} \\
& \ge 4 |Q(x)| \cdot (4m+1)^{s^2m/2} && \text{(since $r \ge 1 + \frac{1}{2} \log_2(4m+1)$)} \\
& \ge 2 |Q(x)| \cdot m^{s^2m/2} \; .
\end{align*}
It thus remains to prove that
\begin{align}
Q(x) & \ge \frac{1}{2} && \text{for all} \ x \in \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \ge \gamma \right\} \; , \\
Q(x) & \le -\frac{1}{2} && \text{for all} \ x \in \bigcup_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\} \; .
\end{align}
